Capstone Movielens Report

### 2025-05-30 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd

  - [calc_day_general_effect](#func.calc_day_general_effect);
  - [calc_day_general_effect.cv](#func.calc_day_general_effect.cv).
:::

@@

::: { #func.UMGY_SmoothedDay_effect.predict.params.day_smoothed_effect .sidebar }

###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

@@

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.params.day_smoothed_effect .sidebar }

###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

@@

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.cv.params.day_smoothed_effect .sidebar }

###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

@@

##### Return
\
The _Root Mean Squared Error (RMSE)_ value of the _UMGYDE Model_ computed as the square root of the _5-Fold Cross Validation MSE_ result.

----------------------------------------------------

@@

@@

::: {.noteblock data-latex=""}
The [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L128) function described below are defined in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L127) section of the [UM-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L128) script.
:::

##### [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#128) Function {#func.regularize.test_lambda.UMGYD_effect.cv}
\

This function calculates _RMSE_ of the _UMGYDE Model_ using _5-Fold Cross Validation_ for the given $\lambda$ parameter value:
```{r, eval=FALSE}
```

>>>>

@@

::: { #func.UMGY_SmoothedDay_effect.predict.params.day_smoothed_effect .sidebar }
###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

@@

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.params.day_smoothed_effect .sidebar }
###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

@@

::: { #func.calc_UMGY_SmoothedDay_effect.MSE.cv.params.day_smoothed_effect .sidebar }
###### ***day_smoothed_effect***
\
A data object representing the _UMGYD Effect_, which can be returned by any of the following functions described above:

@@

##### Return
\
The _Root Mean Squared Error (RMSE)_ of the _UMGYD Model_ computed as the square root of the _5-Fold Cross Validation MSE_ result.

@@

### UMGYDE Model Tuning
\

#### UMGYDE Model Tuning: Support Functions
\

##### [train_UMGY_SmoothedDay_effect.RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L171) Function {#func.train_UMGY_SmoothedDay_effect.RMSE.cv}
\
The function is designed to tune the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function's parameters ([degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) and [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.span)), which is ultimately called internally under the hood.

###### Parameters
\

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree .sidebar }
  - ***degree:***
A parameter to pass to the internally called [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.details) section below) and, ultimately under the hood, to the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) to specify the degree of the polynomials to be used.

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.span .sidebar }
  - ***span:***
A parameter to pass to the internally called [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.details) section below) and, ultimately under the hood, to the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) to control the degree of smoothing.

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.details .sidebar }
###### Details
\
Under the hood, the function internally calls the [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv) and then [calc_UMGY_SmoothedDay_effect.RMSE.cv](#func.calc_UMGY_SmoothedDay_effect.RMSE.cv) described above to compute the `RMSE` of the model to be tuned.
:::

###### Source Code
\
Below is a simplified version of the function's source code:
```{r eval=FALSE}
train_UMGY_SmoothedDay_effect.RMSE.cv <- function(degree = NA, span = NA) {
  train_UMGY_SmoothedDay_effect.cv(degree, span) |>
    calc_UMGY_SmoothedDay_effect.RMSE.cv()
}
```

::: {.noteblock data-latex=""}
The complete version of the source code of the [train_UMGY_SmoothedDay_effect.RMSE.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L171) function is available in the [Tuning `loess` Parameters](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L170) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L171) script on _GitHub_.
:::

###### Return
\
The _Root Mean Squared Error (RMSE)_ of the _UMGYD Model_.

##### [train_UMGY_SmoothedDay_effect.RMSE.cv.degree0](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) Function {#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0}
\
The overloaded version of the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) function described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 0.

###### Parameters
\

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0.params.span .sidebar }
  - ***span:***
A parameter to pass to the internally called [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0.details) section below) and, ultimately under the hood, to the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) to control the degree of smoothing.

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0.details .sidebar }
###### Details
\
Under the hood, the function internally calls the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 0 and the [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.span) parameter value taken from the [parameter of the same name](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0.params.span) of this function to compute the `RMSE` of the model to be tuned.
:::

###### Source Code
\
Below is the function's source code:
```{r eval=FALSE}
train_UMGY_SmoothedDay_effect.RMSE.cv.degree0 <- function(span) {
  train_UMGY_SmoothedDay_effect.RMSE.cv(degree = 0, span)
}
```

::: {.noteblock data-latex=""}
The source code of the [train_UMGY_SmoothedDay_effect.RMSE.cv.degree0](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) function is also available in the [Tuning `loess` Parameters](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L170) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) script on _GitHub_.
:::

###### Return
\
_Root Mean Square Error (RMSE)_ of the _UMGYD_ model trained using the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function called with the given [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree0.params.span) parameter and the `degree` parameter value equal to 0.

##### [train_UMGY_SmoothedDay_effect.RMSE.cv.degree1](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L193) Function {#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree1}
\
The overloaded version of the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) function described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 1.

###### Parameters
\

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree1.params.span .sidebar }
  - ***span:***
A parameter to pass to the internally called [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree1.details) section below) and, ultimately under the hood, to the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) to control the degree of smoothing.

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree1.details .sidebar }
###### Details
\
Under the hood, the function internally calls the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 1 and the [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.span) parameter value taken from the [parameter of the same name](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree1.params.span) of this function to compute the `RMSE` of the model to be tuned.
:::

###### Source Code
\
Below is the function's source code:
```{r eval=FALSE}
train_UMGY_SmoothedDay_effect.RMSE.cv.degree1 <- function(span) {
  train_UMGY_SmoothedDay_effect.RMSE.cv(degree = 1, span)
}
```

::: {.noteblock data-latex=""}
The source code of the [train_UMGY_SmoothedDay_effect.RMSE.cv.degree1](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L193) function is also available in the [Tuning `loess` Parameters](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L170) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L193) script on _GitHub_.
:::

###### Return
\
_Root Mean Square Error (RMSE)_ of the _UMGYD_ model trained using the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function called with the given [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree1.params.span) parameter and the `degree` parameter value equal to 1.

##### [train_UMGY_SmoothedDay_effect.RMSE.cv.degree2](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L196) Function {#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree2}
\
The overloaded version of the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) function described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 2.

###### Parameters
\

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree2.params.span .sidebar }
  - ***span:***
A parameter to pass to the internally called [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) function (see the [Details](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree2.details) section below) and, ultimately under the hood, to the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) to control the degree of smoothing.

::: { #func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree2.details .sidebar }
###### Details
\
Under the hood, the function internally calls the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 2 and the [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.span) parameter value taken from the [parameter of the same name](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree2.params.span) of this function to compute the `RMSE` of the model to be tuned.
:::

###### Source Code
\
Below is the function's source code:
```{r eval=FALSE}
train_UMGY_SmoothedDay_effect.RMSE.cv.degree2 <- function(span) {
  train_UMGY_SmoothedDay_effect.RMSE.cv(degree = 2, span)
}
```

::: {.noteblock data-latex=""}
The source code of the [train_UMGY_SmoothedDay_effect.RMSE.cv.degree2](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L196) function is also available in the [Tuning `loess` Parameters](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L170) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L196) script on _GitHub_.
:::

###### Return
\
_Root Mean Square Error (RMSE)_ of the _UMGYD_ model trained using the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function called with the given [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.degree2.params.span) parameter and the `degree` parameter value of 2.

@@

##### [regularize.train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) Function {#func.regularize.train_UMGYD_effect}
\

This is a helper function that calculates the _Smoothed Day Effect_ of the user ratings and is intended to be used in other functions and scripts (such as [regularize.train_UMGYD_effect](#func.regularize.train_UMGYD_effect)) described below.

###### Parameters
\

::: { #func..params. .sidebar }
  - ***train_set:***
The _Train Set_ dataset.

::: { #func..params. .sidebar }
  - ***lambda:***
As explained [above](#sec.umgydem.support_functions.note.lambda) in the [] section, this is the $\lambda$ coefficient of the penalty of the _Regularization_ method.

::: { #func.regularize.train_UMGYD_effect.details .sidebar }
###### Details
\
Under the hood, the function internally calls the 
:::

###### Source Code
\

The following is the source code of the function:
```{r eval=FALSE}
regularize.train_UMGYD_effect <- function(train_set, lambda) {
  best_degree <- lss.UMGYDE.best_params["degree"]
  best_span <- lss.UMGYDE.best_params["span"]

  train_set |>
    train_UMGY_SmoothedDay_effect(best_degree, 
                                  best_span, 
                                  lambda)
}
```

::: {.noteblock data-latex=""}
The source code of the [regularize.train_UMGYD_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) function is also available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L97) script on _GitHub_.
:::

###### Return
\
A data frame object representing the _User+Movie+Genre+Year+(Smoothed)Day (UMGYD) Effect_.

##### [regularize.train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) Function {#func.regularize.train_UMGYD_effect.cv}
\
The overloaded version of the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) function described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 0.

###### Parameters
\

::: { #func.regularize.train_UMGYD_effect.cv.params.lambda .sidebar }
  - ***lambda:***
As explained [above](#sec.umgydem.support_functions.note.lambda) in the [] section, this is the $\lambda$ coefficient of the penalty of the _Regularization_ method.

::: { #func.regularize.train_UMGYD_effect.cv.details .sidebar }
###### Details
\
Under the hood, the function internally calls the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) described above with the fixed [degree](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.degree) parameter value of 0 and the [span](#func.train_UMGY_SmoothedDay_effect.RMSE.cv.params.span) parameter value taken from the [parameter of the same name](#func.regularize.train_UMGYD_effect.cv.params.span) of this function to compute the `RMSE` of the model to be tuned.
:::

###### Source Code
\
Below is the function's source code:
```{r eval=FALSE}
regularize.train_UMGYD_effect.cv <- function(lambda) {
  best_degree <- lss.UMGYDE.best_params["degree"]
  best_span <- lss.UMGYDE.best_params["span"]

  train_UMGY_SmoothedDay_effect.cv(best_degree, 
                                   best_span, 
                                   lambda)
}
```

::: {.noteblock data-latex=""}
The source code of the [regularize.train_UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) function is also available in the [Tuning `loess` Parameters](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L170) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L190) script on _GitHub_.
:::

###### Return
\
_Root Mean Square Error (RMSE)_ of the _UMGYD_ model trained using the [stats::loess](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/loess) function called with the given [span](#func.regularize.train_UMGYD_effect.cv.params.span) parameter and the `degree` parameter value equal to 0.

##### [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L240) Function {#func.regularize.test_lambda.UMGYD_effect.cv}
\

::: {.noteblock data-latex=""}
The [regularize.test_lambda.UMGYD_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L240) function are defined in the [Regularization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L222) section of the [UM-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L240) script on _GitHub.
:::

This function calculates _RMSE_ of the _UMGYDE Model_ using _5-Fold Cross Validation_ for the given $\lambda$ parameter value:

```{r, eval=FALSE}
regularize.test_lambda.UMGYD_effect.cv <- function(lambda){
  if (is.na(lambda)) {
    stop("Function: regularize.test_lambda.UMGY_effect.cv
`lambda` is `NA`")
  }

  regularize.train_UMGYD_effect.cv(lambda) |>
    calc_UMGY_SmoothedDay_effect.RMSE.cv()
}
```
